Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems

### 2020-09-29 2020/OPTIMA/slides/slides.tex
OPTIMA: fixes, update slides

\begin{frame}
  \frametitle{Problem statement}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{displaymath}
        \begin{array}{cr}\\
          \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
          D=\{y\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i}, \\ 1\leqslant{i}\leqslant{N},\:g_j(y)\leqslant 0, 1\leqslant j\leqslant m\}
        \end{array}
      \end{displaymath}
      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
      \begin{displaymath}
        \begin{array}{cr}\\
          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
        \end{array}
      \end{displaymath}
      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Problem statement}
  Let's consider a set of \(q\) global optimization problems with nonlinear constraints:
  \begin{displaymath}
    \min\left\{\varphi_1(y), y\in D_1 \right\}, \min\left\{\varphi_2(y), y\in D_2\right\},..., \min\left\{\varphi_q(y), y\in D_q\right\}.
  \end{displaymath}
  For an instance, such sets of problems can arise in the following cases:
  \begin{itemize}
    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
    \item a set of problems, obtained after scalarization of a multi-objective problem.
  \end{itemize}
\textbf{Possible approaches}:
  \begin{itemize}
    \item solve each problem independently;
    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
    prioritizing one of the problems.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Index method of handling inequality constraints}
\end{frame}

\begin{frame}
  \frametitle{Пример решения многокритериальной задачи}
  Рассматриваемая задача:
  \begin{displaymath}
    \begin{array}{l}
        Minimize \left \{
        \begin{array}{l}
          f_1(y) = 4 y_1^2 + 4 y_2^2 \\
          f_2(y) = (y_1-5)^2 + (y_2-5)^2 \\
        \end{array}
        \right .
        y_1\in [-1;2],y_2\in [-2;1]
        \\s.t.
        \\
        \left \{
        \begin{array}{l}
          g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
          g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
        \end{array}
        \right .
    \end{array}
  \end{displaymath}
  После использования свёртки Гермейера для скаляризации задача примет вид:
  \begin{displaymath}
    \varphi(y^*(\lambda_1,\lambda_2))=\min_{y\in D}\max\{\lambda_1 f_1(y), \lambda_2 f_2(y)\};\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1
\end{displaymath}
Для численного построения множества Парето выберем
100 наборов коэффициентов \((\lambda_1,\lambda_2)\) таких, что
\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
\end{frame}

\begin{frame}
  \frametitle{Пример решения многокритериальной задачи}
  \vspace*{-1.0cm}
  \begin{displaymath}
    \begin{array}{cr}\\
      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},\:\overline{d}=mean\{d_i\}, \\
      d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
    \end{array}
  \end{displaymath}

  \begin{figure}[ht]
      \centering
      \vspace*{-0.5cm}
      \subfloat[Раздельное решение задач, \(SP_{single}=0.984\)]{
      \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{single_mco.pdf}}}
      \subfloat[Решение множества задач, \(SP_{multi}=0.749\)]{
      \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{multi_mco.pdf}}}
      \caption{Численные оценки множества Парето, полученные после 2500 испытаний}
      \label{fig:mco_pareto}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Тестовые многомерные задачи с ограничениями}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      Тестовые наборы были получены с помощью системы GCGen, позволяющей
      сторить многомерные задачи с ограничениями из заданных функций, и нескольких
      известных генераторов тестовых задач (GKLS, \(F_{GR}\)).
      Характеристики каждого из сгенерированных наборов задач:
      \begin{itemize}
        \item \(q\)=100;
        \item Размерность 2 или 3;
        \item Базовые функции только GKLS или \(F_{GR}\), или их комбинация;
        \item Количество ограничений: 2.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
    \end{column}
  \end{columns}
  \footnotesize{Система GCGen доступна по ссылке \textit{https://github.com/UNN-ITMM-Software/GCGen}}
\end{frame}

\begin{frame}
  \frametitle{Программное и техническое обеспечение}
  \begin{center}
    Реализация параллельного метода была выполнена на языке C++ с использованием технологии OpenMP
    для распареллеливания процесса проведения испытаний на общей памяти.

    Все вычислительные
    эксперименты проведены на машине со следующей конфигурацией: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Результаты экспериментов на наборах синтетических задач}
  \begin{figure}[ht]
    \vspace*{-0.5cm}
      \centering
      \subfloat[\(D_{max}\)]{
      \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_max.pdf}}}
      \subfloat[\(D_{avg}\)]{
      \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_avg.pdf}}}
      \caption{Динамика величин средней по задачам и максимальной точности в процессе решения множества двухмерных задач,
      порождённых двумя разными генераторами GKLS и \(F_{GR}\)}
      \label{fig:devs_mixed}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Результаты экспериментов на наборах синтетических задач}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Класс задач & \textit{p} & Количество итераций & Время, с & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS \& \(F_{GR}\)-based \
        & 1 & 51434 & 90.20 & -    & - \\
        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
      \hline
      GKLS-based 2d \
        & 1 & 59066 & 97.53 & -    & - \\
        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
      \hline
      GKLS-based 3d \
        & 1 & 782544 & 1117.55 & -    & - \\
        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}
  \frametitle{Заключение}
  В ходе работы были достигнуты следующие результаты:
    \begin{itemize}
      \item реализована поддержка нелинейных ограничений в алгоритме, решающeм
      множество задач глобальной оптимизации;
      \item проведены численные эксперименты, демонстрирующие преимущество рассматриваемого подхода в скорости сходимости на всём множестве задач в среднем над решением задач по отдельности;
      \item показана эффективность совместного решения множества задач на примере решения многокритериальной задачи с нелинейными ограничениями.
    \end{itemize}

    В ходе дальнейшей работы планируется:
    \begin{itemize}
      \item улучшить текущую реализацию алгоритма, сократив расходы на содержание поисковой информации для множества задач и тем самым улучшив
      показатели параллельного ускорения по времени;
      \item реализовать версию рассматриваемого алгоритма, работающего на распределенной памяти.
    \end{itemize}
\end{frame}

{
\unnumbered
\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    \Large{Контакты:}
\vspace{0.5cm}

>>>>

\begin{frame}
  \frametitle{Problem statement}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{displaymath}
        \begin{array}{cr}\\
          \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y:\\
           g_j(y)\leqslant 0, 1\leqslant j\leqslant m\},
        \end{array}
      \end{displaymath}
      where \(D\) is a hyperrectangle,
      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
      \begin{displaymath}
        \begin{array}{cr}\\
          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
        \end{array}
      \end{displaymath}
      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Problem statement}
  Let's consider a set of \(q\) global optimization problems with nonlinear constraints:
  \begin{equation}
    \label{eq:many_problems}
    \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
  \min\left\{\varphi_q(y), y\in G_q\right\}.
  \end{equation}
  This work aims to formulate a method that will ensure the uniform convergence to the
  solutions of all problems in the series:
  \begin{equation}
    \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
      \frac{\Vert \overline{y_i}^*(s) - y_i^*\Vert_\infty}{\Vert \overline{y_j}^*(s) - y_j^*\Vert_	\infty} \leqslant \varepsilon,
  \end{equation}
  where \(s\) is the number of step of an optimization method, \(\overline{y_i}^*(s)\) is the best iteration point
  in problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).

\end{frame}

\begin{frame}
  \frametitle{Problem statement}
  Such sets of problems can arise in the following cases:
  \begin{itemize}
    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
    \item a set of problems, obtained after scalarization of a multi-objective problem.
  \end{itemize}
\textbf{Possible approaches}:
  \begin{itemize}
    \item solve each problem independently;
    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
    prioritizing one of the problems.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Index method of handling inequality constraints}
  \begin{enumerate}
    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  \begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
    the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
    Then solving all of the \(q\) problems by MIAGS will
    generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
    \begin{displaymath}
      \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
    \end{displaymath}
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{An example of solving a multi-objective problem}
  Consider the following problem:
  \begin{displaymath}
    \begin{array}{l}
        Minimize \left \{
        \begin{array}{l}
          f_1(y) = 4 y_1^2 + 4 y_2^2 \\
          f_2(y) = (y_1-5)^2 + (y_2-5)^2 \\
        \end{array}
        \right .
        y_1\in [-1;2],y_2\in [-2;1]
        \\s.t.
        \\
        \left \{
        \begin{array}{l}
          g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
          g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
        \end{array}
        \right .
    \end{array}
  \end{displaymath}
  After scalarization the problem has the following form:
  \begin{displaymath}
    \varphi(y^*(\lambda_1,\lambda_2))=\min_{y\in D}\max\{\lambda_1 f_1(y), \lambda_2 f_2(y)\};\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1
\end{displaymath}
To build the Pareto set numerically, consider
100 pairs of the variables \((\lambda_1,\lambda_2)\) such that
\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
\end{frame}

\begin{frame}
  \frametitle{An example of solving a multi-objective problem}
  \vspace*{-1.0cm}
  \begin{displaymath}
    \begin{array}{cr}\\
      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},\:\overline{d}=mean\{d_i\}, \\
      d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
    \end{array}
  \end{displaymath}

  \begin{figure}[ht]
      \centering
      \vspace*{-0.5cm}
      \subfloat[IAGS, \(SP_{single}=0.984\)]{
      \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{single_mco.pdf}}}
      \subfloat[MIAGS, \(SP_{multi}=0.749\)]{
      \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{multi_mco.pdf}}}
      \caption{Numerical estimations of the Pareto set, obtained after 2500 trails}
      \label{fig:mco_pareto}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Generated test problems}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      Sets of test problems are obtained by applying the GCGen system to
      functions produced by generators of unconstrained test problems.
      The main features of the generated sets:
      \begin{itemize}
        \item \(q\)=100;
        \item Dimension 2, 3 or 4;
        \item Base fuctions GKLS or \(F_{GR}\), or a combination of them;
        \item Number of nonlinear constraints: 2.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
    \end{column}
  \end{columns}
  \footnotesize{GCGen is available here \textit{https://github.com/UNN-ITMM-Software/GCGen}}
\end{frame}

\begin{frame}
  \frametitle{Hardware and software environment}
  \begin{center}
    The parallel method was implemented on C++ language with the use of OpenMP
    technology for organizing parallelism on shared memory.

    All the numerical experiments were carried out on a computer with the following
    configuration: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{figure}[ht]
    \vspace*{-0.5cm}
      \centering
      \subfloat[\(D_{max}\)]{
      \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_max.pdf}}}
      \subfloat[\(D_{avg}\)]{
      \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_avg.pdf}}}
      \caption{Dynamic of the average and maximal errors, during solving of problems
      generated on different types of base functions: GKLS и \(F_{GR}\)}
      \label{fig:devs_mixed}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS \& \(F_{GR}\)-based \
        & 1 & 51434 & 90.20 & -    & - \\
        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
      \hline
      GKLS-based 2d \
        & 1 & 59066 & 97.53 & -    & - \\
        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS-based 3d \
        & 1 & 782544 & 1117.55 & -    & - \\
        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
      \hline
      GKLS-based 4d \
        & 1 & 14021720 & 15806.6 & -    & - \\
        & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
        & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
        & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Conclusions}
    \begin{itemize}
      \item implemented support of nonlinear functional constraints in the method
      for solving a set of global optimization problems;
      \item convergence of the implemented method was theoretically proven;
      \item conducted numerical experiments on sets, demonstrating uniform convergence (in the mentioned sense) and
      parallel speedup;
      \item demonstrated efficiency of the implemented method in case of solving a multi-objective problem.
    \end{itemize}
\end{frame}

{
\unnumbered
\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    %\Large{Контакты:}
\vspace{0.5cm}
